Gender and Mental Depression: Analyzing Census Data Through Statistical Methods

### 2024-10-23 paper/paper.qmd
Results and discussion section added

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

Our results are summarized in @tbl-lrmodelresults.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-lrmodelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
# mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
# 
# modelsummary(
#   list(
#     "Mental Illness Exist" = mental_illness
#   ),
#   statistic = "mad"
#   )
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

@@

Weaknesses and next steps should also be included.

@@

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
# 
# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

>>>>

We leverage the logistic regression model to identify the association between suffering from psychological distress with sex and education. We set our model following the @tellingstories guidelines with the rstanarm package [@rstanarm_goodrich, @rstanarm_brilleman] in R [@citer].

Logistic regression is a commonly used regression model when the outcome variable is binary [@tellingstories]. For our dataset, we introduce an outcome variable, which represents whether an individual is suffering from psychological distress. As a predictor variable for the logistic regression, we will use the sex and education variables discussed in the @sec-data.

@@

For logistic regression, we are using the equation below:

\begin{align}
y_i \sim Bern(\pi_i) \\
logit(y_i) \sim \beta_0 + \beta_1 \times sex_i + \beta_2 \times education_i \\
\beta_0 \sim Normal(0,2.5) \\
\beta_1 \sim Normal(0,2.5) \\
\beta_2 \sim Normal(0,2.5)
\end{align}

Here, $y_i$ represents whether a respondent suffers from psychological distress. $sex_i$ represents the respondent's gender, and $education_i$ represents the maximum education attained by the respondent. We randomly select 5000 samples from our entire dataset to simplify the model for lower execution time.

@@

Our results are summarized in @tbl-lrmodelresults-2018 and @tbl-lrmodelresults-2021.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-lrmodelresults-2018
#| tbl-cap: "Model Results for Year 2018"

mental_illness_2018 <- readRDS(file = here::here("models/mental_illness_2018.rds"))

modelsummary(
  list(
    "Severe Mental Illness" = mental_illness_2018
  ),
  statistic = "mad"
)
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-lrmodelresults-2021
#| tbl-cap: "Model Results for Year 2021"

mental_illness_2021 <- readRDS(file = here::here("models/mental_illness_2021.rds"))
modelsummary(
  list(
    "Severe Mental Illness" = mental_illness_2021
  ),
  statistic = "mad"
)
```

@@

## Gender Differences in Depression {#sec-first-point}

@tbl-lrmodelresults-2018 and @tbl-lrmodelresults-2021 demonstrate that the odds of being female and having psychological depression are higher compared to males. The findings are consistent both before and after the pandemic.

## Educational Differences

We have not found any notable differences between education and psychological distress. Research indicates a strong connection between emotional well-being and education, as emotional learning is believed to be nurtured within the school environment and through the curriculum [@norwich_critical_2022;@billington_conditions_2022].

@@

We randomly selected 5000 samples for our model. We have not used any prediction to find out how good our model is for detecting psychological depression. We also have not conducted any data augmentation to improve our model result. Our next steps are carefully curating the dataset, introducing several data validation techniques, and increasing the sample set.

@@
